Temperature and moisture effects on greenhouse gas emissions from deep active-layer boreal soils

### 2016-11-23 manuscript/response_to_reviewer_comments.tex

\medskip
{\it 1.15. Line 238-240: Confusingly worded sentence. }

This has been clarified. {\bf See new lines 264-266.}

\medskip
{\it 1.16. I don't think that the summary of nearby respiration studies add very much to the discussion section. Perhaps if the similarities and discrepancies were more integral to the central findings of the paper or integrated differently into the discussion they would seem more meaningful here. Perhaps comparing with other boreal incubations (eg: Lee et al., 2012; Lavoie et al., 2011) would help to provide some additional context. }

We agree that the comparison to other boreal respiration studies needs improvement, and this echoes Referee \#2's comment 2.4. First, although we think the paragraph about nearby studies provides useful context, we have tightened it considerably. Second, we have restructured and improved the subsequent comparison section, discussing a variety of studies suggested by all the reviewers to better put our results in the context of previous work. {\bf See new lines 295-307 for nearby studies, and expanded context in lines 308-337, 368-372, 401-408, 427-432.}

@@

\medskip
{\it 1.21. Line 322-332: This section, which lays out the crux of the paper, the interaction between temperature and moisture sensitivity in driving microbial respiration is relatively vague. It would be good to describe the less-temperature-sensitive processes that would be important to consider for more stable-C metabolism. And how does moisture play a role here? Perhaps DOC becomes more limiting in the drought conditions? }

This is interesting to consider: what mechanisms might produce a Q10 {\it increase} under drought conditions? This is opposite to what is usually observed (e.g. Jassal et al. 2008), but the field is rife with contradictory results (von Lützow and Kögel-Knabner 2009). We have made this paragraph more specific in this area. {\bf See new lines 378-390.}

\medskip
{\it 1.22. Line 356: The Janssens et al., 2010, citation refers to a meta-analysis of temperate forest soils that are not nitrogen limited. There are studies focusing on Arctic N cycling that would be more appropriate, and many Arctic studies have shown that N availability can limit C mineralization rates. Is this site considered to be N limited in the deep active layer? }

We agree that referring and comparing to studies such as Lavoie et al. (2011) and Bouskill et al. (2014), which focused specifically on high latitudes, would be a useful addition. We don't know of any studies examining the N limitation of deep soils at this site. We now discuss these results in our revision, along with other studies examining the relationship between N availability and C mineralization. {\bf See new lines 392-408.}

@@

We have added caveats, noting in particular the useful but incremental nature of this study. {\bf See new lines 449-460.}

% ======================================================================================
\newpage
\subsection*{Response to Referee \#3}

{\it 3.2. My main criticism is that I think that the authors over-emphasize the results of the daily emissions and that the authors should further explore (or report) the results of the controls of the cumulative C emissions. I'm curious as to whether the relationships with soil C/N and \%N observed in daily emissions still hold on cumulative emissions. The comparison between these soil parameters (i.e. ones that probably don't change much throughout the course of the incubation, including temperature) and the cumulative fluxes is perhaps more appropriate. Perhaps modelers find the controls on daily fluxes interesting and these are likely quite useful in regards to the relationship between moisture and C production (i.e. changes on a daily basis), but I think that the controls on cumulative fluxes are quite interesting and could be further explored. }

We agree that rebalancing the manuscript, focusing a bit more on controls on cumulative emissions and a bit less on the instantaneous fluxes, would strengthen it. Accordingly, we now more fully explore controls on the cumulative emissions, and have moved the table summarizing the instantaneous CH4 flux model, to an appendix. {\bf See new lines 826-832.}

@@

\medskip
{\it 3.3. For example, how do the results of soil properties vs. emissions compare to those of Schädel et al. (2014) and Schädel et al. (2016)? How do the moisture results compare to those of Wickland et al. (2008)? }

The other referees both mentioned this as well. The fact that we didn't cite the Schädel et al. (2016) meta-analysis was a quirk of timing, as it appeared after our manuscript was submitted. We have significantly expanded this, discussing and comparing to Schädel et al. (2014, 2016) and Wickland et al. (2008). {\bf See new lines 46, 81, 313, 368-371, 411, and line 306 for Wickland.}

@@

\medskip
{\it 3.4. I do think that the time series of fluxes could be moved to the supplemental materials if the cumulative fluxes are explored in greater detail. I think this paper could be shortened a little bit although I didn't find the length of the paper onerous. Along these lines, I think that the results summarized above from the cumulative emissions should be included in the abstract. }

We have moved one table to supplementary material (see response to comment 3.2 above), and now summarize cumulative emissions results in the abstract. {\bf See new lines 27-31.}

### 2016-11-23 manuscript/cpcrw_manuscript.Rmd
Final changes, esp. line numbers in response doc

---
output:
  html_document: default
  word_document: default
---
<!---
This Rmarkdown file generates the Biogeosciences manuscript
Built using rmarkdown 1.1, RStudio 1.0.44, and R 3.3.1
November 23, 2016
-->

@@

Interestingly, Q~10~ values were lower in the drought treatment cores, a mathematical consequence of the fact that drought restricted CO~2~ respiration more at 4 °C than at 20 °C. There is evidence that climate warming changes the microbial decay dynamics of soil organic C compounds generally considered to be stable {Frey, 2013 #4062}{Bond-Lamberty, 2016 #4371}. Conditions such as drought can change the amount and quality of DOC available to microbes {e.g., \Lundquist, 1999 #4506}, but we observed no DOC changes between treatments here. Deep active layer soils store large quantities of soil C {Mueller, 2015 #4435} but are not subject to abundant inputs of fresh C from vegetation, so the starting quality of the native soil C in active layer soils is older, more microbially processed, and dominated by more stable "heavy" organic C {Karlsson, 2011 #4436}. Thus, it may not be surprising that these more stable C compounds would be metabolized by processes that have been reported to be less temperature-sensitive.

@@

**Table 1.** Summary of dissolved organic carbon (DOC), percent C, percent N, and CO~2~ and CH~4~ fluxes by treatment. The "Field moisture" and "Drought" columns summarize (mean ± s.d.) 12 cores, combining two groups of N=6 at each incubation temperature, while the "Controlled drought" and "Pre-incubation" columns are N=6.

>>>>

---
output:
  word_document: default
  html_document: default
---
<!---
This Rmarkdown file generates the Biogeosciences manuscript
Built using rmarkdown 1.1, RStudio 1.0.44, and R 3.3.1
November 23, 2016
-->

@@

Interestingly, Q~10~ values were lower in the drought treatment cores, a mathematical consequence of the fact that drought restricted CO~2~ respiration more at 4 °C than at 20 °C. There is evidence that climate warming changes the microbial decay dynamics of soil organic C compounds generally considered to be stable {Frey, 2013 #4062}{Bond-Lamberty, 2016 #4371}. Conditions such as drought can change the amount and quality of DOC available to microbes {\Lundquist, 1999 #4506}, but we observed no DOC changes between treatments here. Deep active layer soils store large quantities of soil C {Mueller, 2015 #4435} but are not subject to abundant inputs of fresh C from vegetation, so the starting quality of the native soil C in active layer soils is older, more microbially processed, and dominated by more stable "heavy" organic C {Karlsson, 2011 #4436}. Thus, it may not be surprising that these more stable C compounds would be metabolized by processes that have been reported to be less temperature-sensitive.

@@

**Table 1.** Summary of dissolved organic carbon (DOC), percent C, percent N, bulk density (BD), and CO~2~ and CH~4~ fluxes by treatment. The "Field moisture" and "Drought" columns summarize (mean ± s.d.) 12 cores, combining two groups of N=6 at each incubation temperature, while the "Controlled drought" and "Pre-incubation" columns are N=6.
